---
title: "DEA_Technical Appendix"
author: "Jordan Gropper"
date: '2022-07-11'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, include = FALSE}
library(gtrendsR)
library(fixest)
library(ggplot2)
library(tidyverse)
library(vtable)
library(Ecdat)
library(ggstance)
library(multcomp)
library(NHANES)
library(purrr)
library(lubridate)
library(corrplot)
library(inspectdf) 
library(dplyr)
```

```{r}
final_df_clean <- read_csv('Lab3_Rawdata/CleanedBase.csv')

```



```{r, eval=FALSE}
#CORRELATION PLOT: 
cor_school <- final_df_clean %>% 
  dplyr::select(CITY, STABBR, HCM2, CONTROL, PBI, ANNHI, TRIBAL, AANAPII, HSI, NANTI, MENONLY, RELAFFIL, SAT_AVG, PCTPELL,
                RET_FT4, PCTFLOAN, month) %>% 
  dplyr::select(where(is.numeric))

#filter relevant items


corrplot(cor(cor_school),
         order = "original",
         diag = FALSE,
         method = "number",
         addCoef.col = 'black',
         type = "upper",
         tl.srt = 45,
         tl.col = "black")

```

```{r}
#TYPE DIAL, Not very useful
final_df_clean %>% 
  inspect_types() %>% 
  show_plot()

#runs
```


```{r}
#HISTOGRAMS:
#index distribution
final_df_clean %>% ggplot(mapping = aes(index)) + 
  geom_histogram(bins = 100, fill="blue")
#average is close to 0
```

```{r}
#SAT Distribution in Bachelors colleges:
final_df_clean %>% ggplot(mapping = aes(SAT_AVG)) + 
  geom_histogram(bins = 10, binwidth = 1000, stat="count", fill ='blue')

#earnings Distributions:
final_df_clean %>% ggplot(mapping = aes(earnings)) + 
  geom_histogram(bins = 100, stat="count", fill ='blue')
# we have a few serious outliers, with the majority of earnings being reported around the $40k 
#range

```

```{r}
final_df_clean %>% ggplot(mapping = aes(factor(year))) + 
  geom_histogram(bins = 100, stat="count", fill ='blue')
#Interesting to see that there aren't as many schools observed in 2016

final_df_clean %>% group_by(year,month) %>% ggplot(mapping = aes(factor(month))) + 
  geom_histogram(bins = 100, stat="count", fill ='blue')
#It seems that the first and last observed months have the only difference in count of observations

```

```{r}
final_df_clean %>% drop_na() %>%  ggplot(mapping = aes(high_earnings)) + 
  geom_histogram(bins = 10, binwidth = 1000, stat="count", fill ='blue')
#the majority of schools in each year aren't high earning

```


```{r}
grouped_control <- final_df_clean %>% group_by(CONTROL) %>% summarize(earnings=mean(earnings, na.rm=TRUE))

grouped_control %>% 
  ggplot(mapping = aes(x=CONTROL, y = earnings)) + 
  geom_col(fill= 'blue') 

```

```{r}
#paterns over months
grouped_month <-  final_df_clean %>% group_by(month) %>% summarize(index = sum(index, na.rm=TRUE))

grouped_month %>% 
ggplot(mapping = aes(x=month, y =index)) + 
  geom_col()

#Grouped by Year:

grouped_year <-  final_df_clean %>% group_by(year) %>% summarize(index = sum(index, na.rm=TRUE))

grouped_year %>% 
  ggplot(mapping = aes(x=year, y =index)) + 
  geom_col()
#it appears that the mean of the indexes of interest were averaging lower in 2016.


```

```{r}


#type of urbanization of the school
final_df_clean %>% group_by(LOCALE) %>%  summarize(m_earnings = mean(earnings, na.rm=TRUE)) %>% 
ggplot(mapping = aes(x=as.factor(LOCALE), y= m_earnings)) + 
  geom_col() 


final_df_clean %>% ggplot(mapping = aes(x=year, y= earnings, fill=as.factor(high_earnings))) + 
  geom_col(position='dodge')

```

## Scatter Plots and Line Graphs:
## Top Schools
I need help with these, and I want to add on a lot more:
```{r}
#trying to do just the top 8 earning school
top_school <- final_df_clean %>% 
  group_by(schname) %>% 
  slice(1) %>% 
  ungroup() %>% 
  arrange(-earnings) %>% 
  head(8)


top_school_data <- final_df_clean %>% filter(unitid %in% top_school$unitid)


top_school_data %>% 
  ggplot(mapping = aes(x = monthorweek, y = index, color = as.factor(schname))) + 
  geom_line()
```


## Bottom Earning Schools:
```{r}
#now do bottom earning schools:
bot_school <- final_df_clean %>% 
  group_by(schname) %>% 
  slice(1) %>% 
  ungroup() %>% 
  arrange(earnings) %>% 
  head(8)


bot_school_data <- final_df_clean %>% filter(unitid %in% bot_school$unitid)

bot_school_data %>% 
  ggplot(mapping = aes(x = monthorweek, y = index, color = as.factor(schname))) + 
  geom_line()
```

```{r}
final_df_clean %>% group_by(high_earnings, monthorweek) %>%  filter(high_earnings != 'NA') %>% 
  summarise(index = mean(index)) %>% 
  ggplot(mapping = aes(x = monthorweek, y = index, color = as.factor(high_earnings))) + 
  geom_line()


```


## Great Box Plots:
These work great!
```{r}

final_df_clean %>% 
  ggplot(aes(x = factor(month), y = index)) +
  geom_boxplot() +
  theme_bw() # GREAT NEW BOXPLOT!


final_df_clean %>% 
  ggplot(aes(x = factor(year), y = index)) +
  geom_boxplot() +
  theme_bw() # GREAT NEW BOXPLOT!



```


```{r}
final_df_clean %>% 
  ggplot(aes(x = factor(high_earnings), y = index)) +
  geom_boxplot() +
  theme_bw()

final_df_clean %>% 
  ggplot(aes(x = factor(score_avail), y = earnings)) +
  geom_boxplot() +
  theme_bw()
```




Would like some visuals that look at the "flag" variables, and if any of those are high earnings as a group, or high interest as a group.
```{r}
#close but not working quite yet!
flag_table <- final_df_clean %>% 
  dplyr::select(PBI, ANNHI, TRIBAL, AANAPII, HSI, NANTI, MENONLY) %>% 
  summarize(across(.fns = sum))


flag_table %>% pivot_longer(cols=everything()) %>%  
ggplot(mapping = aes(x=name, y=value)) + 
  geom_col()
```
HSI Stands for Hispanic Students Flag, meaning that the school is prominetly hispanic. This visual portrays that the highest sum of interest tends to be in Hispanic schools. 









